---
title: "Case Study 2"
author: "Alex Lopez"
date: "11/29/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
# YouTube Presentation Link: https://youtu.be/zKjWktTrOm0

# Introduction
In this analysis we will be reviewing and interpreting variables and characteristics of employee Attrition, or turn over. We will examine data with variables such as, wage, satisfaction poll scores, employee department, and so on. The objectives of the analysis will be to determine the 3 top factors contributing to Attrition, report on job role specific trends, identify other key insight, and finally, build prediction models for attrition and wage for employees that quit.

```{r setup, include=FALSE, fig.width=6,fig.asp=0.618}
knitr::opts_chunk$set(echo = TRUE)
# Package Load-In
rm(list = ls())
library(tidyverse)
library(readr)
library(corrplot)
library(ggplot2)
library(caret)
library(e1071)
library(class)
library(fpp)
library(GGally)
library(lsr)
library(ggcorrplot)
library(fastDummies)
library(Metrics)
```

# Data Load-In
```{r}
# Load original data set.
orig_data <- as.data.frame(read_csv(url("https://raw.githubusercontent.com/alexanderlopez1793/CaseStudy2DDS/main/CaseStudy2-data.csv")))
```

# Data Pre-Processing
```{r}
# review data set
str(orig_data)
summary(orig_data)
# checking for missing values not captured in summary()
sum(is.null(orig_data))
# convert Attrition to factor for comparison
orig_data$Attrition = as.factor(orig_data$Attrition)
# convert chr variables to factors since none have high value count
orig_data$BusinessTravel = as.factor(orig_data$BusinessTravel)
orig_data$Department = as.factor(orig_data$Department)
orig_data$EducationField = as.factor(orig_data$EducationField)
orig_data$Gender = as.factor(orig_data$Gender)
orig_data$JobRole = as.factor(orig_data$JobRole)
# may group single and divorced together
orig_data$MaritalStatus = as.factor(orig_data$MaritalStatus)
orig_data$OverTime = as.factor(orig_data$OverTime)
# dropping variables with no beneficial info
# no value in employee ID #
mod_data = subset(orig_data, select = -c(ID))
# Employee # also has no value
mod_data = subset(mod_data, select = -c(EmployeeNumber))
# count is the same for every employee
mod_data = subset(mod_data, select = -c(EmployeeCount))
# standard hours the same for all employee
mod_data = subset(mod_data, select = -c(StandardHours))
# all employees over 18, will be dropped
mod_data = subset(mod_data, select = -c(Over18) )
```

# EDA - Univariate Categorical
```{r}

# review Categorical variables
mod_data %>% ggplot(aes(x=Attrition)) + geom_bar(fill = "blue",col = "black") + labs(title="Attrition Bar Plot")  + geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=BusinessTravel)) + geom_bar(fill = "blue",col = "black") + labs(title="Business Travel Bar Plot")+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=Department)) + geom_bar(fill = "blue",col = "black") + labs(title="Department Bar Plot")+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=EducationField)) + geom_bar(fill = "blue",col = "black") + labs(title="Education Field Bar Plot") + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=Gender)) + geom_bar(fill = "blue",col = "black") + labs(title="Gender Bar Plot")+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=JobRole)) + geom_bar(fill = "blue",col = "black") + labs(title="Job Role Bar Plot")+ theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=MaritalStatus)) + geom_bar(fill = "blue",col = "black") + labs(title="Marital Status Bar Plot")+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
mod_data %>% ggplot(aes(x=OverTime)) + geom_bar(fill = "blue",col = "black") + labs(title="Overtime Bar Plot")+ geom_text(stat='count', aes(label=..count..), vjust=-.2)
```

## Insights: 

- Attrition: Data is somewhat biased with a majority (84%) of employees showing "No" to Attrition (turnover).
- Business Travel: Most employees (71%) rarely travel.
- Department: Almost 2/3 (65%) of employees work in "Research & Development".
- Education Field: Employees with "Life Sciences" and "Medical" education backgrounds make up almost 3/4 of data (72%).
- Gender: Employees are 59% male.
- Job Role: 60% of employees work as "Sales Executive", "Research Scientist", or "Laboratory Technician".
- Marital Status: There is almost a 50/50 split between "Married" and "Single"/"Divorced" employees ("Divorced" employees can be considered single as well).
- Overtime: 71% of employees do not work overtime.

# EDA - Univariate Numerical
```{r}
# review Numeric variables
mod_data %>% ggplot(aes(x=Age)) + geom_histogram(binwidth=5,fill = "blue",col = "black")+ labs(title="Age Histogram Plot")
mod_data %>% ggplot(aes(x=DailyRate)) + geom_histogram(binwidth=100,fill = "blue",col = "black")+ labs(title="Daily Rate Histogram Plot")
mod_data %>% ggplot(aes(x=DistanceFromHome)) + geom_histogram(binwidth=2,fill = "blue",col = "black")+ labs(title="Distance From Home Histogram Plot")
mod_data %>% ggplot(aes(x=Education)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Education Histogram Plot")
mod_data %>% ggplot(aes(x=EnvironmentSatisfaction)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Environment Satisfaction Histogram Plot")
mod_data %>% ggplot(aes(x=HourlyRate)) + geom_histogram(binwidth=5,fill = "blue",col = "black")+ labs(title="Hourly Rate Histogram Plot")
mod_data %>% ggplot(aes(x=JobInvolvement)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Job Involvement Histogram Plot")
mod_data %>% ggplot(aes(x=JobLevel)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Job Level Histogram Plot")
mod_data %>% ggplot(aes(x=JobSatisfaction)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Job Satisfaction Histogram Plot")
mod_data %>% ggplot(aes(x=MonthlyIncome)) + geom_histogram(binwidth=1000,fill = "blue",col = "black")+ labs(title="Monthly Income Histogram Plot")
mod_data %>% ggplot(aes(x=MonthlyRate)) + geom_histogram(binwidth=1000,fill = "blue",col = "black")+ labs(title="Monthly Rate Histogram Plot")
mod_data %>% ggplot(aes(x=NumCompaniesWorked)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Number of Companies Worked Histogram Plot")
mod_data %>% ggplot(aes(x=PercentSalaryHike)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Percent Salary Hike Histogram Plot")
mod_data %>% ggplot(aes(x=PerformanceRating)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Performance Rating Histogram Plot")
mod_data %>% ggplot(aes(x=RelationshipSatisfaction)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Relationship Satisfaction Histogram Plot")
mod_data %>% ggplot(aes(x=StockOptionLevel)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Stock Option Level Histogram Plot")
mod_data %>% ggplot(aes(x=TotalWorkingYears)) + geom_histogram(binwidth=2,fill = "blue",col = "black")+ labs(title="Total Working Years Histogram Plot")
mod_data %>% ggplot(aes(x=TrainingTimesLastYear)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Training Time Last Year Histogram Plot")
mod_data %>% ggplot(aes(x=WorkLifeBalance)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Work/Life Balance Histogram Plot")
mod_data %>% ggplot(aes(x=YearsAtCompany)) + geom_histogram(binwidth=2,fill = "blue",col = "black")+ labs(title="Years at Company Histogram Plot")
mod_data %>% ggplot(aes(x=YearsInCurrentRole)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Years in Current Role Histogram Plot")
mod_data %>% ggplot(aes(x=YearsSinceLastPromotion)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Years Since Last Promotion Histogram Plot")
mod_data %>% ggplot(aes(x=YearsWithCurrManager)) + geom_histogram(binwidth=1,fill = "blue",col = "black")+ labs(title="Years With Current Manager Histogram Plot")
```

## Insights: 

- Age: Employee age has approx. normally distribution.
- Hourly/Daily/Monthly Rate: There is no obvious statistical distribution.
- Distance From Home: There is a right skew with a large amount of employees living very close to work.
- Education: There is a some-what normal distribution around mid-level education (assumption is 5 is the highest level of education i.e. PHD, and 1 is equal to little or no post high school education).
- Environment Satisfaction: A majority of employees have high work satisfaction (3 or 4 rating).
- Job Involvement: Most employees have mid-range job involvement (assumption is 1 is low involvement and 4 is high involvement).
- Job Level: There is a right skew with a majority of employees having lower level jobs.
- Job Satisfaction: A majority of employees have high levels of job satisfaction.
- Monthly Income: There is a right skew with, what seems to be, positive outliers (high wages).
- Number of Companies Worked: There is a right skew, with a large number of employees having only worked 1 other job.
- Percent Salary Hike: There is a right skew, with a majority of employees recieving low salary percentage increases.
- Performance Rating: Scale is likely from 1-5, meaning all employees show to have mid to mid-high ratings.
- Relationship Satisfaction: A majority of employees have high relationship satisfaction with there supervisor/manager (assumption is 1 is low satisfaction and 4 is high satisfaction).
- Stock Option: Most employees have low level stock option plan, likely correlating to high number of employees with lower Job Level.
- Total Working Years: There is a right skew, with positive outliers appearing to be present.
- Training Total Last Year: Most employees only trained 2-3 times last year.
- Work/Life Balance: A majority of employees show to have a good level of work/life balance (assumption 1 is bad work/life balance and 4 is good work/life balance).
- Years at Company: There is right skew, likely correlates to number of total years worked.
- Years in Current Role: Some what right skew, could also be bi-modal around 2 years and 7 years.
- Years Since Last Promotion: There is a right skew, may have strong correlation to years in current role.
- Years With Current Manager: Some what right skew, could also be bi-modal. Distribution looks very similar to years in current role.

# EDA - Bivariate Categorical
```{r}
#business travel
mod_data %>%
  group_by(BusinessTravel) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = BusinessTravel, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - Business Travel") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =10)) + geom_text(aes(label=attrition_rate), size = 10, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))
# department
mod_data %>%
  group_by(Department) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Department, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - Department") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =10)) + geom_text(aes(label=attrition_rate), size = 10, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))
# education field
mod_data %>%
  group_by(EducationField) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = EducationField, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - Education Field") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =10, angle = 60,hjust = 1)) + geom_text(aes(label=attrition_rate), size = 5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 30))
# gender
mod_data %>%
  group_by(Gender) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = Gender, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - Gender") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =10)) + geom_text(aes(label=attrition_rate), size = 10, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 20))
# Job Role
mod_data %>%
  group_by(JobRole) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = JobRole, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - JobRole") + theme(plot.title = element_text(size =15),axis.text.x = element_text(size =9, angle = 60,hjust = 1)) + geom_text(aes(label=attrition_rate), size = 5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 50))
# Marital Status
mod_data %>%
  group_by(MaritalStatus) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = MaritalStatus, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - MaritalStatus") + theme(plot.title = element_text(size =15),axis.text.x = element_text(size =10)) + geom_text(aes(label=attrition_rate), size = 5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 50))
# Overtime
mod_data %>%
  group_by(OverTime) %>%
  summarise(attrition_rate = round((sum(if_else(Attrition == "Yes",1,0))/n()*100),2)) %>%
  ggplot(aes(x = OverTime, y = attrition_rate))+ geom_bar(stat = 'identity',fill = "blue") + ggtitle("Attrition Rate - OverTime") + theme(plot.title = element_text(size =15),axis.text.x = element_text(size =10)) + geom_text(aes(label=attrition_rate), size = 5, position=position_dodge(width=0.2), vjust=-0.25)+ scale_y_continuous(limits = c(0, 50))
```

## Insights:

- Business Travel: Highest rate of Attrition is for employees who travel often.
- Department: Highest rate of Attrition is for employees who work in sales.
- Education Field: Attrition rates are highest amoungst Human Resources and Technical Degrees.
- Gender: Attrition rate is only just larger for males.
- Job Role: Attrition rate is almost half for Sales Rep's. Also, Attrition for Human Resources, Lab Tech, and Research Scientist is around 20%.
- Marital Status: Attrition rate is over 25% for single employees.
- Overtime: Attrition rate for employees that work overtime is over 30%.

# EDA - Bivariate Numeric
```{r}
# Age
mod_data %>%
  ggplot(aes(x = Age, fill = Attrition)) + geom_boxplot() + ggtitle("Age by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Hourly rate
mod_data %>%
  ggplot(aes(x = HourlyRate, fill = Attrition)) + geom_density(alpha = .5) + ggtitle("Hourly Rate by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Daily rate
mod_data %>%
  ggplot(aes(x = DailyRate, fill = Attrition)) + geom_density(alpha = .5) + ggtitle("Daily Rate by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Monthly rate
mod_data %>%
  ggplot(aes(x = MonthlyRate, fill = Attrition)) + geom_density(alpha = .5) + ggtitle("Monthly Rate by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Distance From Home
mod_data %>%
  ggplot(aes(x = DistanceFromHome, fill = Attrition)) + geom_density(alpha = .5) + ggtitle("Distance From Home by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Education
mod_data %>%
  ggplot(aes(x = Education, fill = Attrition)) + geom_bar() + ggtitle("Education by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Environment Satisfaction
mod_data %>%
  ggplot(aes(x = EnvironmentSatisfaction, fill = Attrition)) +  geom_bar() + ggtitle("Environment Satisfaction by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Job Involvement
mod_data %>%
  ggplot(aes(x = JobInvolvement, fill = Attrition)) +  geom_bar() + ggtitle("Job Involvement by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Job Level
mod_data %>%
  ggplot(aes(x = JobLevel, fill = Attrition)) +  geom_bar() + ggtitle("Job Level by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Job Satisfaction
mod_data %>%
  ggplot(aes(x = JobSatisfaction, fill = Attrition)) +  geom_bar() + ggtitle("Job Satisfaction by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Monthly Income
mod_data %>%
  ggplot(aes(x = MonthlyIncome, fill = Attrition)) +  geom_boxplot() + ggtitle("Monthly Income by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Number of Companies Worked
mod_data %>%
  ggplot(aes(x = NumCompaniesWorked, fill = Attrition)) +  geom_bar() + ggtitle("Number of Companies Worked by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Percent Salary Hike
mod_data %>%
  ggplot(aes(x = PercentSalaryHike, fill = Attrition)) +  geom_bar() + ggtitle("Percent Salary Hike by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Performance Rating
mod_data %>%
  ggplot(aes(x = PerformanceRating, fill = Attrition)) +  geom_bar() + ggtitle("Performance Rating by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Relationship Satisfaction
mod_data %>%
  ggplot(aes(x = RelationshipSatisfaction, fill = Attrition)) +  geom_bar() + ggtitle("Relationship Satisfaction by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Stock Option
mod_data %>%
  ggplot(aes(x = StockOptionLevel, fill = Attrition)) +  geom_bar() + ggtitle("Stock Option by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Total Working Years
mod_data %>%
  ggplot(aes(x = TotalWorkingYears, fill = Attrition)) +  geom_boxplot() + ggtitle("Total Working Years by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Training Total Last Year
mod_data %>%
  ggplot(aes(x = TrainingTimesLastYear, fill = Attrition)) +  geom_bar() + ggtitle("Training Total Last Year by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Work/Life Balance
mod_data %>%
  ggplot(aes(x = WorkLifeBalance, fill = Attrition)) +  geom_bar() + ggtitle("Work/Life Balance by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Years at Company
mod_data %>%
  ggplot(aes(x = YearsAtCompany, fill = Attrition)) +  geom_boxplot() + ggtitle("Years at Company by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Years in Current Role
mod_data %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) +  geom_boxplot() + ggtitle("Years in Current Role by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Years Since Last Promotion
mod_data %>%
  ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) +  geom_boxplot() + ggtitle("Years Since Last Promotion by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
# Years With Current Manager
mod_data %>%
  ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) +  geom_boxplot() + ggtitle("Years With Current Manager by Attrition") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =10))
```

## Insights:

- Age: over 75% of Employees that quit were under 40.
- Hourly/Daily/Monthly Rate: No noticeable trends, when comparing rates and carrying them over to days and then months, the Attrition values do not equal each other.
- Distance From Home: Attrition rate was higher for those employees who did not travel far for work.
- Education: Attrition rate is highest among employees with lowest score.
- Environment Satisfaction: Attrition rates are highest among employees with the lowest environment satisfaction score.
- Job Involvement: Attrition rates are highest among employees with lowest job involvement.
- Job Level: Attrition rates are highest in lowest job level, but rates follow same overall distribution.
- Job Satisfaction: Attrition rates are highest among employees with low job satisfaction score.
- Monthly Income: 75% of employees who quit had monthly incomes of 6000 or less.
- Number of Companies Worked: Attrition rate was highest among employees who had only worked one job before, but that may be biased since a that subgroup is a strong majority.
- Percent Salary Hike: Attrition rates were highest among employees with the lowest salary increases by percentage.
- Performance Rating: Most employees had performance ratings of 3 so the increased attrition rate is expected.
- Relationship Satisfaction: Attrition rates are highest among employees with low relationship satisfaction.
- Stock Option: Employees with the lowest stock option plan, had the highest attrition rate.
- Total Working Years: 75% of the employees that quit, had 10 or less working years.
- Training Total Last Year: Attrition rates appear to follow same overall distribution for training times.
- Work/Life Balance: Attrition rate is highest for employees with lowest Work/Life Balance.
- Years at Company: Over 75% of the employees that quit, had been with the company for less than 9 years.
- Years in Current Role: Over 75% of the employees that quit, had only been in their current role for less than 5 years.
- Years Since Last Promotion: Over 75% of the employees that quit, had just been promoted in the last 3 years.
- Years With Current Manager: Over 75% of the employees that quit, had worked with their last manager for 6 years or less.

# Top 3 Factors
```{r}
# find strongest correlations with Attrition and plot correlation matrix
corr_data1 = model.matrix(~0+., data=mod_data) %>% cor(use="pairwise.complete.obs")
corr_data1 = as.data.frame(corr_data1)
corr_data1 %>% filter(AttritionYes> .07) %>% ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=4, show.legend = TRUE, title = "Correlation Matrix")
# overtime,job role (sales rep), marital status (single)
```

## Insights: 

The Top 3 Factors for Attrition are listed below:

- Overtime (Corr = 0.27): Whether or not an employee works overtime has the highest correlation to Attrition. Overtime is work outside of the standard 40 hour work week. Being over worked as an employee can certainly lead to distraught feelings and eventually, voluntary turnover. 
- Job Role, Sales Representative (Corr = 0.20): Job Role, more specifically, Sales Representative has the second highest correlation to Attrition. Looking at the bi-variate analysis, sales rep employees had the highest attrition rate at 45%. This percentage is amplified when considering sales reps make up only 6% of the total employee population. 
- Marital Status, Single (Corr = 0.18): Single Employees had the third highest correlation to Attrition. One thought process can be, employees with no family are more inclined to take risks, such as quitting to become entrepreneurs or moving to another company.


# Job Specific Trends
```{r}
# using ggpairs and group_by tables to review job specific trends

# initial review of numeric variables of interest
mod_data %>% select(JobRole,MonthlyIncome,PercentSalaryHike) %>% ggpairs(legend=1,aes(color=JobRole))

mod_data %>% group_by(JobRole) %>% summarise(meanMI = mean(MonthlyIncome),count=n())
mod_data %>% group_by(JobRole) %>% summarise(meanPSH = mean(PercentSalaryHike),count=n())

# initial review of categorical variables of interest
mod_data %>% select(JobRole,EnvironmentSatisfaction,JobSatisfaction,MonthlyIncome) %>% ggpairs(legend=1,aes(color=JobRole))

mod_data %>% select(JobRole,MaritalStatus,OverTime) %>% ggpairs(legend=1,aes(color=JobRole))
mod_data %>% select(JobRole,Education,MonthlyIncome,OverTime) %>% ggpairs(legend=1,aes(color=JobRole))

mod_data %>% select(JobRole,Education,MonthlyIncome) %>% ggpairs(legend=1,aes(color=JobRole))
mod_data %>% select(JobRole,JobSatisfaction,DistanceFromHome) %>% ggpairs(legend=1,aes(color=JobRole))

mod_data %>% group_by(JobRole) %>% 
  summarise(meanYAC = mean(YearsAtCompany),
            meanES=mean(EnvironmentSatisfaction),
            meanJS=mean(JobSatisfaction),
            meanMI = mean(MonthlyIncome),
            meanE = mean(Education),
            meanDFH = mean(DistanceFromHome),
            meanRS = mean(RelationshipSatisfaction),
            count=n())
```

## Insights:

- Sales Representatives have the lowest mean monthly income (2653), lowest mean education level (2.45) and are the least experienced employees with a mean of 2.9 years with the company.
- Mangers and Research Directors have the highest mean monthly incomes (17197 and 15750 respectively) and are the most experienced employees with 13.8 and 10.2 years with the company.
- Healthcare Representatives have the highest mean job satisfaction score (2.83), even with the highest mean distance from home (9.8).

# Other Interesting Trends
```{r}
# trends and other insights using visuals

mod_data %>% 
  filter(Attrition == "Yes") %>%
  select(JobRole,OverTime,MaritalStatus) %>% 
  ggpairs(legend=1,aes(color=JobRole))

mod_data %>% 
  ggplot(aes(y=MonthlyIncome,x=as.factor(JobSatisfaction),fill=Attrition)) + geom_boxplot() + labs(title = "Attrition by Job Satisfaction vs Monthly Income",y="Monthly Income",x="Job Satisfaction") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =9)) + facet_wrap(~JobRole)

```

## Insights

- In general, jobs with higher wages tend to have better employee retention. In this case, job satisfaction did not appear have significant impact on attrition - Managers and Directors saw very little attrition regardless of satisfaction score. Overall, this plot provides evidence that if the company wants to decrease attrition, the best course of action will likely be to increase wages.

# Classifcation Model
```{r}
# set seed for repeatability
set.seed(1)
# create dummy variables for factor variables
dum_data = dummy_cols(mod_data,select_columns = c('BusinessTravel','Department','EducationField','Gender','JobRole','MaritalStatus','OverTime'),remove_selected_columns = TRUE)

# scale all original numeric variables with large ranges
# dum_data$Age = as.numeric(scale(dum_data$Age))
# dum_data$DailyRate = as.numeric(scale(dum_data$DailyRate))
# dum_data$DistanceFromHome = as.numeric(scale(dum_data$DistanceFromHome))
# dum_data$HourlyRate = as.numeric(scale(dum_data$HourlyRate))
# dum_data$MonthlyIncome = as.numeric(scale(dum_data$MonthlyIncome))
# dum_data$MonthlyRate = as.numeric(scale(dum_data$MonthlyRate))
# dum_data$NumCompaniesWorked = as.numeric(scale(dum_data$NumCompaniesWorked))
# dum_data$PercentSalaryHike = as.numeric(scale(dum_data$PercentSalaryHike))
# dum_data$TotalWorkingYears = as.numeric(scale(dum_data$TotalWorkingYears))
# dum_data$TrainingTimesLastYear = as.numeric(scale(dum_data$TrainingTimesLastYear))
# dum_data$YearsAtCompany = as.numeric(scale(dum_data$YearsAtCompany))
# dum_data$YearsInCurrentRole = as.numeric(scale(dum_data$YearsInCurrentRole))
# dum_data$YearsSinceLastPromotion = as.numeric(scale(dum_data$YearsSinceLastPromotion))
# dum_data$YearsWithCurrManager = as.numeric(scale(dum_data$YearsWithCurrManager))

# create train and test sets
trainInd = sample(seq(1,870,1), .7*870)
train = dum_data[trainInd,]
test = dum_data[-trainInd,]

# will use knn classifier for attrition classification model
classifications = knn(train[,c(1,3:52)],test[,c(1,3:52)],train$Attrition,prob = TRUE, k = 3)
confusionMatrix(table(classifications,test$Attrition))

# trying naive bayes
model = naiveBayes(train[,c(1,3:52)],train$Attrition,laplace = 1)
table(predict(model,test[,c(1,3:52)]),test$Attrition)
CM = confusionMatrix(table(predict(model,test[,c(1,3:52)]),test$Attrition))
CM

```

## Insights:

- KNN Model Results: 
  - Accuracy = 78%
  - Sensitivity = 91%
  - Specificity = 7%
- Naive Bayes Model Results: 
  - Accuracy = 72%
  - Sensitivity = 70%
  - Specificity = 83%
- Naive Bayes model provided required metric scores and so can be used as our Attrition classification model.

# Regression Model
```{r}
# training/test sets have already been created
# creating model - linear regression
model2 = lm(MonthlyIncome ~ Age + YearsAtCompany + JobLevel, data=train)
summary(model2)

# predictions
model2_preds = predict(model2,newdata = test)
#as.data.frame(model2_preds)

# using rmse function, RMSE= 1372.05
rmse(test$MonthlyIncome,model2_preds)

# creating another model - linear regression w/ years at company removed due to high p-value
model3 = lm(MonthlyIncome ~ Age + JobLevel, data=train)
summary(model3)

# predictions
model3_preds = predict(model3,newdata = test)
#as.data.frame(model3_preds)

# using rmse function, RMSE= 1372.05
rmse(test$MonthlyIncome,model3_preds)

# plot for presentation
mod_data %>% ggplot(aes(x=Age,y=MonthlyIncome,color=as.factor(JobLevel))) + geom_point() + labs(title = "Age vs Monthly Income per Job Level",y="Monthly Income",x="Age",color="Job Level") + theme(plot.title = element_text(size =20),axis.text.x = element_text(size =9))

```

## Insights:

- Linear Regression model yields RMSE = 1372.05. Variables considered were selected after reviewing data during EDA. Job Level seems to hold the most significance when predicting monthly income (p-value = <2e-16).

# Conclusion
In conclusion, the top 3 factors for attrition were identified to be over time work, job role (sales rep), and marital status (single). Further analysis and interpretation was also provided for specific job roles like sales rep and managers. Insight on attrition reduction was also provied. Finally, 2 models that meet the given criteria were identified. For attrition classification, the Naive Bayes will provide metric scores over 70% and for a monthly income prediction, the linear regression model will provide an RMSE less than $1375.